One-shot learning: From domain knowledge to action models

### 2019-02-20 doc/ijcai-19-deductivelearning/deductivelearning.tex
Introduction updated

% Use the postscript times font!
\usepackage{times}
\usepackage{soul}
\usepackage{url}
\usepackage[hidelinks]{hyperref}
\usepackage[utf8]{inputenc}
\usepackage[small]{caption}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage{algorithmic}
\urlstyle{same}

@@

@@

\section{Introduction}
\label{sec:introduction}
The specification of action models is a complex process that limits, too often, the application of {\em model-based planning} to real-world tasks~\cite{kambhampati:modellite:AAAI2007}. The {\em machine learning} of action models relieves this {\em knowledge acquisition bottleneck} of {\em model-based planning} and nowadays, there exists a wide range of effective approaches for learning action models~\cite{arora:amodels:ker2018}. Many of the most successful approaches for learning planning action models are however purely {\em inductive}~\cite{yang2007learning,pasula2007learning,mourao2010learning,zhuo2013action}, linking learning performance exclusively to the {\em amount} and {\em quality} of the input learning examples (which typically are observation of plan executions).

This paper addresses the learning of action models exploiting a different source of knowledge, {\em deductive} knowledge. Our approach leverages {\em state-invariants} (i.e. logic formulae that specify constraints about the possible states of a given domain) to cushion the negative impact of insufficient learning examples. Given an action model, state-of-the-art planners infer {\em state-invariants} from that model to reduce the search space and make the planning process more efficient~\cite{helmert2009concise}. In this paper we follow the opposite direction and leverage {\em state-invariants} to learn the planning action model. The benefit of learning action models from {\em state-invariants} is two-fold, {\em state-invariants} constrain the space of possible action models and can complete learning examples that are only partially observed.

Our approach for learning \strips\ action models from {\em state-invariants} is to compile the learning task into a classical planning task. Our compilation is built on top of the classical planning compilation for the learning of \strips\ action models~\cite{aineto2018learning} and it is flexible to different kinds of input knowledge including both partial observations of plan executions and {\em state-invariants}. The compilation outputs an \strips\ action model that is {\em consistent} with all the given input knowledge. The experimental results demonstrate that, even at unfavorable scenarios where input observations are minimal (a single learning example that comprises just a full initial state and a partially observed state), {\em state-invariant} are helpful to learn good quality \strips\ action models.

>>>>

% Use the postscript times font!
\usepackage{times}
\usepackage{soul}
\usepackage{url}
\usepackage[hidelinks]{hyperref}
\usepackage[utf8]{inputenc}
\usepackage[small]{caption}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{xspace}
\urlstyle{same}

@@

\newcommand{\ARMS}{{\small {\sffamily ARMS}}\xspace}
\newcommand{\CAMA}{{\small {\sffamily CAMA}}\xspace}
\newcommand{\SLAF}{{\small {\sffamily SLAF}}\xspace}
\newcommand{\LAMP}{{\small {\sffamily LAMP}}\xspace}
\newcommand{\NOISTA}{{\small {\sffamily NOISTA}}\xspace}
\newcommand{\LOCM}{{\small {\sffamily LOCM}}\xspace}
\newcommand{\LOCMtwo}{{\small {\sffamily LOCM2}}\xspace}
\newcommand{\LOP}{{\small {\sffamily LOP}}\xspace}
\newcommand{\AMAN}{{\small {\sffamily AMAN}}\xspace}
\newcommand{\LOUGA}{{\small {\sffamily LOUGA}}\xspace}

@@

\section{Introduction}
\label{sec:introduction}

The learning of action models in planning has been typically addressed with inductive learning data-intensive approaches. From the pioneer learning system \ARMS~\cite{yang2007learning} to other more recent ones \cite{MouraoZPS12,zhuo2013action,kuvcera2018louga}, all of them require of the order of thousands of actions (input plan observations) to obtain and validate an action model. Generally speaking, the rationale behind these approaches is to obtain the statistically significant model $M$ that best explains the plan observations by minimizing some error and redundancy metrics. Specifically, given a plan observation $o=\langle s_0,a_1,s_1,\ldots,s_{n-1},a_n,s_n\rangle$, being $s_i, i>0$, a possibly partially observable state; if a solution plan $\pi$ to the planning task $\tup{s_0,s_n}$ can be produced with $M$ such that $\pi$ contains the observed actions $\tup{a_1,\ldots,a_n}$ of $o$, and the trace resulting from the execution of $\pi$ comprises the (partially) observed states $\tup{s_1,\ldots,s_{n-1}}$ of $o$, then $M$ explains $o$. Validating the explainability of a model as an optimization task over a test set of observations neither guarantees completeness (the model may not explain all the observations) nor correctness (a state that results from the execution of $\pi$ may contain contradictory information even though the state comprises the observed state).

Differently, other approaches rely on a symbolic-via learning. The Simultaneous Learning and Filtering (\SLAF) approach~\cite{AmirC08} exploits logical inference and builds a complete explanation of observations through a CNF formula that represents the initial belief state, and an observation $o$ that contains partially observable states. The formula is updated with every action and state of $o$, thus representing all possible transition relations consistent with $o$. \SLAF extracts all satisfying models of the learned formula with a SAT solver although the algorithm cannot effectively learn the preconditions of actions. A more recent approach addresses the learning of action models from $o$ as a planning task that searches the space of the all possible action models~\cite{aineto2018learning}. A plan here is conceived as a series of steps that determine the preconditions and effects of the model actions plus other steps that validate the formed actions in $o$. The advantage of this approach is that only requires about a total of 50 actions in the set $o$.

After this brief analysis, we raise the following question: can an action model be learnt from \emph{no observations} but using domain-specific knowledge?. The question is motivated by (a) the assumption that obtaining enough training observations is often difficult and costly, if not impossible in some domains~\cite{Zhuo15}; (b) we may not know the physics of the real-world domain being modeled but we may know certain pieces of knowledge about the domain; and (c) obtaining correct action models that are usable beyond their applicability to a set of testing observations. To this end, we opted for checking our hypothesis in the framework proposed in ~\cite{aineto2018learning} since this planning-based satisfiability approach allows us to configure additional constraints in the compilation scheme, it is able to work under a minimal set of observations and uses an off-the-shelf planner\footnote{We thank authors for providing us with the source files of their learning system.}. Ultimately, we aim to compare the informational power of domain observations against the representational power of domain-specific knowledge.

The paper is organized as follows ...

%The specification of action models is a complex process that limits, too often, the application of {\em model-based planning} to real-world tasks~\cite{kambhampati:modellite:AAAI2007}. The {\em machine learning} of action models relieves this {\em knowledge acquisition bottleneck} of {\em model-based planning} and nowadays, there exists a wide range of effective approaches for learning action models~\cite{arora:amodels:ker2018}. Many of the most successful approaches for learning planning action models are however purely {\em inductive}~\cite{yang2007learning,pasula2007learning,mourao2010learning,zhuo2013action}, linking learning performance exclusively to the {\em amount} and {\em quality} of the input learning examples (which typically are observation of plan executions).

%This paper addresses the learning of action models exploiting a different source of knowledge, {\em deductive} knowledge. Our approach leverages {\em state-invariants} (i.e. logic formulae that specify constraints about the possible states of a given domain) to cushion the negative impact of insufficient learning examples. Given an action model, state-of-the-art planners infer {\em state-invariants} from that model to reduce the search space and make the planning process more efficient~\cite{helmert2009concise}. In this paper we follow the opposite direction and leverage {\em state-invariants} to learn the planning action model. The benefit of learning action models from {\em state-invariants} is two-fold, {\em state-invariants} constrain the space of possible action models and can complete learning examples that are only partially observed.

%Our approach for learning \strips\ action models from {\em state-invariants} is to compile the learning task into a classical planning task. Our compilation is built on top of the classical planning compilation for the learning of \strips\ action models~\cite{aineto2018learning} and it is flexible to different kinds of input knowledge including both partial observations of plan executions and {\em state-invariants}. The compilation outputs an \strips\ action model that is {\em consistent} with all the given input knowledge. The experimental results demonstrate that, even at unfavorable scenarios where input observations are minimal (a single learning example that comprises just a full initial state and a partially observed state), {\em state-invariant} are helpful to learn good quality \strips\ action models.
